Pattern matching for\\ $(213,231)$-avoiding permutations

### 2015-02-06 article.tex
Proof of Proposition 1
modifciation : precision on which subsequence to take to have a match in lemma 2

%Stripe
\DeclareMathOperator{\stripea}{s}
\newcommand{\stripe}[2]{\stripea_{{#1}}({#2})}
\newcommand{\stripew}[1]{\stripea_{{#1}}}

@@

\begin{lemma}
\label{lemma:MatchStripeToPermutation}
	Let $\pi$ and $\sigma$ be two $(213,231)$-avoiding permutations.
	Then $\pi$ contains $\sigma$ if and only if 
	$s_\sigma$ is a subsequence of $s_\pi$.
\end{lemma}

\begin{proof} 
	The forward direction is obvious.
	We prove the backward direction by induction on the size of the motif
	$\sigma$.
	The base case is a motif of size 2. Suppose that we have the motif $(12)$, thus the stripe of the motif is $\ustep$. If there is a match in stripe, then there exists two (contiguous) increasing elements in the text, which is a match in permutation. Same goes for the permutation $(21)$. Now, suppose that the lemma is true for every motif in $\Av{n}{(231,213)}$. Let $\pmotif$ be a motif in $\Av{n+1}{(231,213)}$, and $\ptext$ a permutation avoiding $(213,231)$. Suppose that there is a match of $\stripew{\pmotif}$ in $\stripew{\ptext}$, Let's $s$ be the subsequence such that $\stripew{\pmotif}$ occurs in $\stripew{\ptext}$.
	By induction $\pmotif[2:]$ appears in $\ptext$ by subsequence $s[2:]$.
	Moreover if the first element the stripe of the motif is ascend (descent), then by lemma \ref{lemma:first element is 1 or n} the first element is the minimal (maximal) element. Remark that a subsequence of a permutation avoiding (231,213) is also avoiding (231,213), thus the first element of the subsequence $s$ is either the minimal (maximal), element which complete our proof.
\qed
\end{proof}

We are now ready to solve the pattern containment in case
both $\pi$ and $\sigma$ are $(213, 231)$-avoiding.

@@

%		\subsection{With Motif Avoiding (231,213)}	
%		
%			\textbf{Proposition \num} Let $\pmotif$, $\ptext$ 
%			be two permutation,
%			with  $\pmotif$ a permutation avoiding (231,213), 
%			we can decide in $O(|\pmotif|*|\ptext|^3)$ time
%			and $O(|\ptext|^3) space$
%			if $\pmotif$ 
%			appear in $\ptext$.\\	
%
%			\textbf{Proof}
%			We consider the following problem :\\
%			
%			Given a motif $\pmotif$ avoiding $(213,231)$,
%			a text $\ptext$, $i$,$j$ , $i<j$.\\
%			
%			$\mathbb{P}_{\pmotif,\ptext,min,max}(i,j)= \left\{ 
%				\begin{array}{l l}
%					true & \text{If the sufixe of $\pmotif$ starting at $i$}\\
%						& \text{is a pattern of the suffix of $\ptext$} \\
%						& \text{with every element in $[min,max]$}\\
%		
%					false & otherwise\\
%				\end{array} \right.$\\
%		
%			$\mathbb{P}_{\pmotif,\ptext,min,max}$ is closed under induction. 
%			it can be
%			solved by means of the following relations:\\
%			
%			
%			\textbf{BASE :} \\
%			$\mathbb{P}_{\pmotif,\ptext,min,max}(|\pmotif|,j)= \left\{ 
%					\begin{array}{l l}
%						true & \text{if $min < \ptext(j) < max$
%						}\\
%						false & otherwise\\
%					\end{array} \right. $\\	
%		
%			\textbf{STEP :} \\
%			$\mathbb{P}_{\pmotif,\ptext,min,max}(i,j)= \left\{ 
%					\begin{array}{l l}
%						False & \text{if $\ptext(j)<min$ or $max<\ptext(j)$}\\
%						\bigcup_{j<k}\mathbb{P}_{\pmotif,\ptext,t(j),max}(i+1,k) 
%							& \text{if $\\mathcal{S}(\pmotif,i)=\ustep$
%							}\\
%						\bigcup_{j<k}\mathbb{P}_{\pmotif,\ptext,min,t(j)}(i+1,k) 
%							& \text{if $\\mathcal{S}(\pmotif,i)=\dstep$
%							}\\
%					\end{array} \right. $\\
%			At each step, 
%			we are match the current element of $\pmotif$ ($\pmotif(i)$)
%			with the current element of $\ptext$ ($\ptext(j)$), if possible.
%			then we are match the rest of $\pmotif$ with 
%			every suffixes $\ptext$ starting after $j$.
%			Clearly there is a match if and only if 	$\bigcup_{0<k<|\ptext|,0<min<max<|\ptext|}\mathbb{P}_{\pmotif,\ptext,min,t(j)}(0,k)$
%			is true.   
%			This implies a $O(|\pmotif|*|\ptext|^3)$ time
%			and $O(|\ptext|^3) space$ algorithm.\\

@@

\section{Pattern Matching With Motif Avoiding (231,213)}	
\label{section:sigma only avoid 231 and 213}

>>>>

%Stripe
\DeclareMathOperator{\stripea}{s}
\newcommand{\stripe}[2]{\stripea_{{#1}}[{#2}]}
\newcommand{\stripew}[1]{\stripea_{{#1}}}

@@

%\begin{corollary}
%
%If $\pi \in \Av{n}{(231,213)}$ its stripe can be define such as, \\
%$\stripe{\pi}{i} = \left\{ 
%	\begin{array}{l l}
%			\ustep & \quad \text{if $\pi[i]$ is the minimal element of $\pi[i:]$}\\    
%			\dstep & \quad \text{if $\pi[i]$ is the maximal element of $\pi[i:]$}\\
%\end{array} \right.$ 
%\end{corollary}

The following lemma is central to all our algorithm.

\begin{lemma}
\label{lemma:MatchStripeToPermutation}
Let $\pi$ and $\sigma$ be two $(213,231)$-avoiding permutations. Then $\pi$ contains $\sigma$ by a subsequence $s$ if and only if $s_\sigma$ is a subsequence of $s_\pi$ by the same subsequence $s$.
\end{lemma}

\begin{proof} 
	The forward direction is obvious.
	We prove the backward direction by induction on the size of the motif
	$\sigma$.
	The base case is a motif of size 2. Suppose that we have the motif $(12)$, thus the stripe of the motif is $\ustep$. If there is a match in stripe, then there exists two (contiguous) increasing elements in the text, which is a match in permutation. Same goes for the permutation $(21)$. Now, suppose that the lemma is true for every motif in $\Av{n}{(231,213)}$. Let $\pmotif$ be a motif in $\Av{n+1}{(231,213)}$, and $\ptext$ a permutation avoiding $(213,231)$. Suppose that there is a match of $\stripew{\pmotif}$ in $\stripew{\ptext}$, Let's $s$ be the subsequence such that $\stripew{\pmotif}$ occurs in $\stripew{\ptext}$.
	By induction $\pmotif[2:]$ appears in $\ptext$ by subsequence $s[2:]$.
	Moreover if $\stripe{\pmotif}{1}$ is ascend (descent), then by lemma \ref{lemma:first element is 1 or n} $\pmotif[1]$ is the minimal (maximal) element of the motif. Furthermore $\stripew{\pmotif}$ occurs in $\stripew{\ptext}$ by subsequence $s$, so $\stripe{s}{1}$ is also ascend (descent) thus $s[1]$ is also the minimal (maximal) element of the subsequence $s$ which complete our proof.
\qed
\end{proof}

We are now ready to solve the pattern containment problem in case
both $\pi$ and $\sigma$ are $(213, 231)$-avoiding.

@@

@@

\section{$\sigma$ only is $(231,213)$-avoiding}	
\label{section:sigma only avoid 231 and 213}
